---
title: "Data Tidying"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```




*set warning and messages are set to false so that they do not show up when you knit the document*
```{r, message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(readr)
```

-   only need to install packages once, then you just need to use the library() tool

### Knitting tips

-   knitting is a good way to make sure your code is reproducible
-   it reruns all your code in a fresh environemnt so it will tease out some error

### Reading in a dataset

right-click the download button from the website where the data is located...select "copy link" and then paste the link below.

-   this is good for making data reproducible and sharing it with colleges
-   read_csv (from readr is better for reading in large documents and is a little smarter than read.csv which is a base r package

```{r}
catch_origonal <- read_csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1")
```

## Clean data

%>% is a pipe operator which connects the output of one function to the next argument

-   reads as "and then"

-   good for writing efficient code

    -   BUT breaking things up into intermediate steps helps to make it more readable
    -   strike a balance between the two

-   shortcut is Ctrl + Shift + M

select() function can be used to pull out specifc columns you want, or can be used to drop specific columns

```{r}
catch_data <- catch_origonal %>% 
  #select(Region, Year, Chinook, Coho, Pink, Chum, Sockeye)
  select(-All, -notesRegCode)

#ways to check out what your data looks like
  #head(catch_origonal)
  #glimpse()
  #summary()
```

Chinook column is listed as character when we know it should be numeric...

-   use the code below to try and change the column to numeric

    -   we get an error saying at least one value was changed to NA not a number
    -   so we need to look at the data and see which row(s) gave the error

    ```{r}
    catch_clean <- catch_data %>%
      mutate(Chinook = as.numeric(Chinook))
    ```

Looking for error in the file...

-   which() function returns indices where something is true

```{r}
i <- which(is.na(catch_clean$Chinook))
catch_origonal[i, ] # this is used to print the i row and all the columns
```

Change the character from I to 1 if_else() says if this condition = "something", then replace with "something else"

```{r}
catch_clean <- catch_data %>%
  mutate(Chinook = if_else(Chinook == "I", "1", Chinook)) %>%
  mutate(Chinook = as.numeric(Chinook))
```

### Reshaping data

Changing from wide to long format

**-C()**

-   specifies what columns NOT to pivot over

**names_to =**

-   names the new column that the old ones are condenses into

**values_to =**

-   names the new column that the values are being put into

**rename()**

- easily rename column names 
- and other things??

```{r}
catch_long <- catch_clean %>%
  pivot_longer(cols = -c(Region, Year), names_to = "Species", values_to = "Catch") %>%
  rename(catch_thousands = Catch) %>%
  mutate(catch = catch_thousands * 1000)


head(catch_long)
```

"..." in descriptions means other arguments you can use

- unnamed vectors will be used so you probably will need ""


**starts_with()**

-   selects things based on what they start with

Changing from long to wide
```{r}
catch_wide <- catch_long %>%
  pivot_wider(names_from = Region, values_from = catch)

head(catch_wide)
```

**Quoted and unquoted variable names**

- if a column name already exists, you don't need quotes
- if you are creating a new variable name then you must use quotes
  - there are some inconsistancies where it will let you run functions using quotes areound variable names that already exist 

## Split-Apply-Combine

```{r}
mean_region <- catch_long %>%
  filter(Species == "Sockeye") %>%
  group_by(Region, Species) %>%
  summarise(catch_mean = mean(catch), 
            n = n())

head(mean_region)
```
**group_by()**

- stores other information you cant readily see
- .groups = "drops" drops the grouping again

**summarise()**

- can add n = n() to get the number of observations used 
    - probably good for catching issues with averaging across replicates (ie; for LTER cover data if it averages across 7 quads instead of 8 because it dropped a row due to a 0 value)

**filter()**

- works a lot like "subset()"

## Joins

loading in a second datset with full region names
```{r}
region_defs <- read.csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.303.1") %>% 
    select(code, mgmtArea)

head(region_defs)
```
this gives and error because there are no common column names
```{r}
#catch_joined <- left_join(catch_long, region_defs)
```
The addion to this code specifys wich columsn to join by if the coumn names do not match (instead of having to rename them)

- either way you should still add the "by = " part so that it doesn't just guess what to combine by
    - catch_joined <- left_join(catch_long, region_defs, by = "Region")
```{r}
catch_joined <- left_join(catch_long, region_defs, by = c("Region" = "code"))

nrow(catch_joined) == nrow(catch_long)
```
**anti-join**
  *haven't gone over this but something you can look into*

- shows you what doesn't have matches



















