---
title: "Data visualization"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(leaflet)
  library(dplyr)
  library(tidyr)
  library(ggplot2)
  library(DT)
  library(scales) # install.packages("scales")
```

```{r}
data_url <- "https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3Af119a05b-bbe7-4aea-93c6-85434dcb1c5e"

esc <- tryCatch(
    read.csv("data/escapement.csv"),
    error=function(cond) {
        message(paste("Escapement file does not seem to exist, so get it from the KNB."))
        esc <- read.csv(url(data_url, method = "libcurl"))
        return(esc)
    }
)

head(esc)
```

## TryCatch

- saying to use local file if you have it, if not, print a message and then grab the dataset from on online source
- makes code more robust and reproducible because anyone can use it wether they have the file locally or not


```{r}
annual_esc <- esc %>%
  separate(sampleDate, into = c("Year", "Month", "Day"), sep = "-") %>%
  mutate(Year = as.numeric(Year))

Mean_esc_year <- annual_esc %>%
  group_by(Year, SASAP.Region, Species) %>%
  summarise(escapement = sum(DailyCount), 
            n = n()) %>%
  filter(Species %in% c("Sockeye", "Chinook", "Coho", "Pink", "Chum"))

head(Mean_esc_year)
```

instead of listing [Species == "Sockeye" | Species == "Chinook" ...ect]

- use %in% c() 

**lubridate package** 

- lubridate::year()
- better for dealing with dates


```{r}

ggplot(Mean_esc_year, aes(x = Species, y = escapement, fill = SASAP.Region)) +
  geom_col()+
  scale_color_brewer(palette = "Dark2")+
theme()
```

**create your own theme call**
```{r}
my_theme <- theme_bw()+
  theme(legend.position = "bottom")

Mean_esc_year %>%
  filter(SASAP.Region == "Kodiak") %>% 
ggplot(aes(x = Year, y = escapement, color = Species)) + 
   geom_line() +
    geom_point() +
    scale_y_continuous(labels = comma, n.breaks = 4) +
    ylab("Escapement") +
    ggtitle("Kodiak Salmon Escapement") +
    my_theme
```

**Patchwork Package**

- good for combining distinct graphs into one figure

## Leaflet
- makes interactive maps

```{r}
locations <- esc %>%
  distinct(Location, Latitude, Longitude) %>%
  drop_na

leaflet(locations) %>% 
  addTiles() %>% 
  addMarkers(lng = ~Longitude, lat = ~Latitude, popup = ~ Location)
```






