---
title: "Untitled"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

changing echo = in the set up chuck 

# header 1
## header 2
### header 3
#### header 4

sample text **bold text**

Need to have spaces between the line above to put text into a new paragraph 

* item 1
* item 2
* item 3

*italic text*

1. item 1
2. item 2
3. item 3

or use just "1's" and it will order it automaticaly

1. item 1
1. item 2
1. item 3

Add code chunk with 

+ (ctrl + alt + i)

running code

- cntrl + enter
  - good for running individual expressions/lines of code within a code chunk

```{r}
x <- 3 + 2
x <- 2
```

## Read in data

```{r}
  bg_chem <- read.csv("../Data/BGchem2008data.csv")
```
pressing "Tab" after the ../ autofill options for you to select from 

## Analysis 


## Plots 
```{r}
plot(bg_chem$CTD_Depth, bg_chem$CTD_Temperature)
```
this is the same code as above but will print the graph without showing the code when its knit together 

```{r, echo=FALSE}
plot(bg_chem$CTD_Depth, bg_chem$CTD_Temperature)
```


see:
help>cheat sheets > r markdown cheat sheet (or reference guide) for different chunk options 


# intro to git 

- learned how to connect Rstudio to github, and visa versa 
- learned how to make edits and push to github
    - helpful for version control when working in groups




