---
title: "test_LTER"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(tidyr)
library(readr)
```


```{r}
LTER_cover_origonal <- read_csv("https://portal.edirepository.org/nis/dataviewer?packageid=knb-lter-arc.20084.2&entityid=66dcbe2b73b3f5bb606072705e5f1fd7")
```

```{r}
unique(LTER_cover_origonal$Species)
```

**Check unique vales in each column to make sure that there are not naming errors**
small differences in how the names of a species was abbreviated will be read in as a different species
  Common mistakes include:
  - extra spaces
  - differences in capitalization
  - periods being in different spots 
  - spelling errors
    - eg; `bare` `Bare` `Bare ` and `Bare.` will all be read as a unique species name instead of as one species
```{r}
unique(LTER_cover_origonal$Species)
```

If there are mistakes then rename them using the code below, then recheck unique values again to make sure the recode worked *and you didn't loose any observations*
```{r}

# example of how to correct naming convention errors 
#LTER_cover_origonal$Species <- recode(LTER_cover_origonal$Species, "Bare" = "bare")
```

**Now I want to group some like things togeather that are currently identified to a level more specific than necessary**

- eg. Standing dead 'St. D.' is identified to species level but we don't need it to be so specific for our needs so we will *change the name* to just standing dead evergreen 'St. D. Evrg'
```{r}
#simplify groups 

  #group all standing dead evergreens together
LTER_cover_origonal$Species <- recode(LTER_cover_origonal$Species, "St. D. Led Pal" = "St. D. Evrg")
LTER_cover_origonal$Species <- recode(LTER_cover_origonal$Species, "Dead Evrg" = "St. D. Evrg")
LTER_cover_origonal$Species <- recode(LTER_cover_origonal$Species, "Dead Bet" = "St. D. Evrg")
LTER_cover_origonal$Species <- recode(LTER_cover_origonal$Species, "St. D. Bet" = "St. D. Evrg")

  #group all grasses together 
LTER_cover_origonal$Species <- recode(LTER_cover_origonal$Species, "Cal can" = "grass")
LTER_cover_origonal$Species <- recode(LTER_cover_origonal$Species, "Grass ex." = "grass")
LTER_cover_origonal$Species <- recode(LTER_cover_origonal$Species, "Car big" = "grass")
LTER_cover_origonal$Species <- recode(LTER_cover_origonal$Species, "Arc lat" = "grass")
LTER_cover_origonal$Species <- recode(LTER_cover_origonal$Species, "Cal lap" = "grass")
LTER_cover_origonal$Species <- recode(LTER_cover_origonal$Species, "Fes sp." = "grass")

  #group all St. D together 
LTER_cover_origonal$Species <- recode(LTER_cover_origonal$Species, "St. D. Sal" = "St. D")
LTER_cover_origonal$Species <- recode(LTER_cover_origonal$Species, "St. D. Bet." = "St. D")
LTER_cover_origonal$Species <- recode(LTER_cover_origonal$Species, "St. D. Evrg" = "St. D")  

  #group all litter together 
LTER_cover_origonal$Species <- recode(LTER_cover_origonal$Species, "Eri vag litter" = "litter")

```

**Now you can remove any data that you will not need for the analysis. In this case we do not want to include block 4**

- we do not use block 4 because its not a true replicate of blocks 1-3

```{r}

LTER_cover_origonal <- LTER_cover_origonal[!LTER_cover_origonal$Block == "4", ]

```

**vectors for assigning functional groups**
```{r}

bare <- c("Bare ground")
dec_shrub <- c("Deciduous shrubs")
ev_shrub <- c("Evergreen shrubs")
grass <- c("Grass")
sedge <- c("Sedge")
lichen <- c("Lichen")
litter <- c("Litter")
moss <- c("Moss")
Forb <- c("Forb")
St.D <- c("Standing dead")

```

**Subset the dataset by Species that you want to group into a functional group**

```{r}

  ## Bare ground ##
bg <- subset(LTER_cover_origonal, species== "bare" | species== "frost boil") 

  ## Deciduous shrubs ##
ds <- subset(LTER_cover_origonal, species== "Bet nan" | species== "Vac uli" | species== "Sal pul" | species== "Arc alp" | species== "Sal phl" | species== "Sal arc" | species== "Sal ret")

 ## Evergreen shrubs ##
es <- subset(LTER_cover_origonal, species== "Emp nig" | species== "Led pal" | species== "Vac vit" | species== "Cas tet" | species== "And pol" | species== "Dry int")

 ## Graminoid ##
gr <- subset(LTER_cover_origonal, Species== "grass" | Species== "graminoid")

 ## Sedge ##
sed <- subset(LTER_cover_origonal, Species== "Eri vag")


  ## Forb ##
fb <- subset(LTER_cover_origonal, Species== "Ped lep" | Species== "Ped lap" | Species== "Pol bis" | Species== "Rub cha" | Species== "Luz sp." | Species== "Pet sp." | Species== "Hie alp" | Species== "Ped sp." | Species== "Arnica sp." | Species== "Ste edw" | Species== "Pet fri" | Species== "Ste lon" | Species== "Sau ang" | Species== "Hie sp." | Species== "Min sp." | Species== "Tof sp." | Species== "Ste sp." | Species== "dicot" | Species== "Lag gla" | Species== "Sax pun" | Species== "Pyr sec" | Species== "Equ arv" | Species== "Pol viv" | Species== "Tof coc" | Species== "Pyr sp.")

  ## other groups ##

lich <- subset(LTER_cover_origonal, Species== "lichen")

lit <- subset(LTER_cover_origonal, Species== "litter")
  
ms <- subset(LTER_cover_origonal, Species== "moss")

std <- subset(LTER_cover_origonal, Species== "St. D")


```













!!!!!!!!!!!!!!!!LEFT OFF HERE!!!!!!!!!!!!!!!















```{r}

MyMerge       <- function(x, y){
  df            <- merge(x, y, all = TRUE)
  rownames(df)  <- df$Row.names
  df$Row.names  <- NULL
  return(df)
}
relcov.func_08_21           <- Reduce(MyMerge, list(bg1, bg2, ds1, ds2, ds3, ds4, ds5, ds6, ds7, es1, es2, es3, es4, es5, es6, gram1, gram2, gram3, forb1, forb2, forb3, forb4, forb5, forb6, forb7, forb8, forb9, forb10, forb11, forb12, forb13, forb14, forb15, forb16, forb17, forb18, forb19, forb20, forb21, forb22, forb23, forb24, lichen, litter, ms, standing.d))


#export this file

write.csv(relcov.func_08_21, file = "relcov_func_all.csv")

```

## Statistics & graphing

\#combine renamed groups

```{r}

#first we must sum the relative cover values across function groups within quadrats
sum.func <- (relcov.func_08_21) %>% group_by(year, block, plot, quad, func.group) %>% summarise_at(vars(rel.cov), list(relcov = sum ), na.rm = TRUE)
```

subset into different years based on \# of quads measured per plot

```{r}

Q4 <- subset(sum.func, year== "2008")
Q5 <- subset(sum.func, year== "2015" | year== "2020")
Q8 <- subset(sum.func, year== "2010" | year== "2011" | year== "2012" | year== "2013" | year== "2014" | year== "2016" | year== "2017" | year== "2018" | year== "2019" |year== "2010" |year== "2021")

```

Calculate stats across quads within plots

```{r}

#Sum across quads in a plot
Q4_sum_avg_quad <- (Q4) %>% group_by(year, block, plot, func.group) %>% summarise_at(vars(relcov), list(sum1 = sum), na.rm = TRUE)

Q5_sum_avg_quad <- (Q5) %>% group_by(year, block, plot, func.group) %>% summarise_at(vars(relcov), list(sum1 = sum), na.rm = TRUE)

Q8_sum_avg_quad <- (Q8) %>% group_by(year, block, plot, func.group) %>% summarise_at(vars(relcov), list(sum1 = sum), na.rm = TRUE)

#Vectors for averages
avg.Q4.quad <- c(Q4_sum_avg_quad$sum1/4)
avg.Q5.quad <- c(Q5_sum_avg_quad$sum1/5)
avg.Q8.quad <- c(Q8_sum_avg_quad$sum1/8)

#creates and fills a new column with the average (by diving that sum by the number of reps specified)
Q4_sum_avg_quad["avg.1"] <-avg.Q4.quad
Q5_sum_avg_quad["avg.1"] <-avg.Q5.quad
Q8_sum_avg_quad["avg.1"] <-avg.Q8.quad

#-------------------------------------------------------------------------------------------------------#
MyMerge       <- function(x, y){
  df            <- merge(x, y, all = TRUE)
  rownames(df)  <- df$Row.names
  df$Row.names  <- NULL
  return(df)
}
Avg.func.quad           <- Reduce(MyMerge, list(Q4_sum_avg_quad, Q5_sum_avg_quad, Q8_sum_avg_quad))

#export this file
write.csv(Avg.func.quad, file = "avg_relcov_quad.csv")
```

Calculate stats across blocks

```{r}

Avg.func.block <- (Avg.func.quad) %>% group_by(year, plot, func.group) %>% summarise_at(vars(avg.1), list(avg.2 = mean, block.sd = sd), na.rm = TRUE)

#export this file
write.csv(Avg.func.block, file = "avg_relcov_block.csv")

```

\#\#line graph \~ average rel.cov

```{r}
#specify factor levels for fertilization and clipping treatments 

Avg.func.block$plot <- factor(Avg.func.block$plot, levels = c("CT", "F2", "F5", "F10"))
Avg.func.block$func.group <- factor(Avg.func.block$func.group, levels = c("Deciduous shrubs", 
                                                                          "Evergreen shrubs", 
                                                                          "Eri vag",
                                                                          "Graminoid",
                                                                          "Forb",
                                                                          "Lichen",
                                                                          "Moss", 
                                                                          "Litter",
                                                                          "Standing dead",
                                                                          "Bare ground"))

#Creates stacked bar graph#
ggplot(data = Avg.func.block, aes(fill=func.group, x = year, y = avg.2)) +
  geom_bar(position="fill", stat="identity")+ #Percent stacked
  #geom_bar(position="stack", stat="identity")+ #Stacked
  scale_fill_manual(" ", values = c(
    "Deciduous shrubs" = "#004c6d",
                                                           #"Deciduous shrubs" = "red",
                                                           "Evergreen shrubs" = "#6996b3",
                                                           "Eri vag" = "#036d3f",
                                                           #"Eri vag" = "red",
                                                           "Graminoid"="#6fa17e",
                                                          #"Graminoid"="red",
                                                           "Forb"="#c4d6c8",
                                                          #"Forb"="red",
                                                           "Lichen"="#f6e8c3",
                                                           "Moss"="#dfc27d",
                                                           "Litter"="#bf812d",
                                                           "Standing dead"="#8c510a",
                                                           "Bare ground" = "#543005"))+
  theme_light()+
  theme(aspect.ratio = 9/18.5)+ #use for vertical stack only
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  theme(axis.text.x = element_text(color = "black", size = 8, angle = 45, hjust = 1))+
  theme(axis.text.y = element_text(color = "black", size = 6))+
  theme( axis.line = element_line(colour = "black", 
                      size = 1, linetype = "solid"))+
  scale_y_continuous(breaks= c(0, 0.50, 1))+
  theme(plot.title = element_text(hjust = .5))+
  labs(y = "Mean relative cover", x = " ")+ 
  #theme(axis.title.y = element_text(face = "bold"))+
    #theme(axis.title.x = element_text(face = "bold"))+
  #theme(legend.direction="horizontal")+
  #theme(legend.position="bottom")+
  facet_grid(plot ~ .) #vertically stacked

ggsave("stacked_bar_cover.jpeg")
```
